Total Score

### 2023-03-14 src/MSAtotal_scores.Rmd
renamed student scores to equity scenario

library(tidyverse)
library(janitor)
library(googlesheets4)

@@

climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301", sheet = "climate_risk+EAL") %>% 
  select(cbsa, normalized_inverted_climate_risk, inverted_n_eal_score)

@@

# reorder columns
col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2", "normalized_inverted_climate_risk", "inverted_n_eal_score", "inverted_n_health_emissions_cost")

criteria_all <- criteria_all[, col_order] %>% 
  select(!inverted_n_eal_score)

@@

```{r}
# weighting criteria
weighted_criteria <- criteria_all %>% 
  select(4:10)

weighted <- t(t(weighted_criteria)*criteria_weights_client)
```

@@

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(final_tot_students = cbsa,
         real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score)) #this is just a place holder

size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$normalized_inverted_climate_risk[i]
                  )
  criteria_all$final_tot_students[i] = weighted.mean(cbsa_data, criteria_weights_students)
}

@@

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(client_FINAL_score = cbsa, #this is just a place holder
         score_climate50_prop_EAL = cbsa)

size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$normalized_inverted_climate_risk[i]
                  )
  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
}

@@

```{r rename columns and output to google sheet}

criteria_all <- criteria_all %>% 
  rename(financial_score = normalized_irr,
         co2_e = inverted_n_total_co2,
         health_impact = inverted_n_health_emissions_cost,
         electricity = normalized_annual_renewable,
         climate_risk = normalized_inverted_climate_risk) %>% 
  mutate_if(is.numeric, ~round(., 5))

@@

```{r client data viz}
## read in weighted criteria sheet
client_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen") %>% 
  unite("msa_state", city_msa:state, sep = ", ")

# read in cbsa data with full names
score_tidy <- client_weighted_criteria %>% 
  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
               names_to = "criteria",
               values_to = "criteria_score") %>% 
 filter(msa_state != "Portland, ME") %>%  # removing portland (negative number)
  group_by(cbsa)

@@

# read in cbsa data with full names
score_tidy2 <- student_weighted_criteria %>% 
  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
               names_to = "criteria",
               values_to = "criteria_score") %>% 
# filter(msa_state != "Portland, ME") %>%  # removing portland (negative number)
  group_by(cbsa)

>>>>

library(tidyverse)
library(janitor)
library(googlesheets4)
library(dplyr)
```

@@

climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301", sheet = "climate_risk+EAL") %>% 
  select(cbsa, normalized_inverted_climate_risk)

@@

# reorder columns
col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2",  "inverted_n_health_emissions_cost", "normalized_inverted_climate_risk")

criteria_all <- criteria_all[, col_order] %>% 
  rename(financial_score = normalized_irr,
         co2_e_score = inverted_n_total_co2,
         health_impact_score = inverted_n_health_emissions_cost,
         electricity_score = normalized_annual_renewable,
         climate_risk_score = normalized_inverted_climate_risk) %>% 
  mutate(real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score)) %>% 
  mutate_if(is.numeric, ~round(., 5))

@@

@@

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(equity_scenario_score = cbsa,
         client_FINAL_score = cbsa) #this is just a place holder

for (i in 1:nrow(criteria_all)) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$financial_score[i],
                 criteria_all$co2_e_score[i],
                criteria_all$health_impact_score[i],
                criteria_all$electricity_score[i],
                criteria_all$climate_risk_score[i]
                  )
  criteria_all$equity_scenario_score[i] = weighted.mean(cbsa_data, criteria_weights_students)
}

@@

# multiply column by respective weight and sum for total

for (i in 1:nrow(criteria_all)) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$financial_score[i],
                 criteria_all$co2_e_score[i],
                criteria_all$health_impact_score[i],
                criteria_all$electricity_score[i],
                criteria_all$climate_risk_score[i]
                  )
  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
}

@@

```{output to google sheet}

criteria_all <- criteria_all %>% 
  mutate_if(is.numeric, ~round(., 5))

weighted <- t(t(weighted_criteria)*criteria_weights_client)

@@

```{r client data viz}
## read in weighted criteria sheet
client_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen") %>% 
  unite("msa_state", city_msa:state, sep = ", ") 
 # mutate(test2 = rowSums(client_weighted_criteria[,c(3:9)]))

# read in cbsa data with full names
score_tidy <- client_weighted_criteria %>% 
  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
               names_to = "criteria",
               values_to = "criteria_score") %>% 
 filter(msa_state != "Portland, ME")  # removing portland (negative number)

@@

# read in cbsa data with full names
score_tidy2 <- student_weighted_criteria %>% 
  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
               names_to = "criteria",
               values_to = "criteria_score")
